A 30 Million Year Old Mini-Neptune in the Kepler Field

### 2021-05-27 paper/ms.tex
update ms

\documentclass[12pt,twocolumn,tighten]{aastex63}
%\documentclass[12pt,twocolumn,tighten,linenumbers]{aastex63}
%\documentclass[12pt,twocolumn,tighten,trackchanges]{aastex63}
\usepackage{amsmath,amstext,amssymb}
\usepackage[T1]{fontenc}
\usepackage{apjfonts}
\usepackage[figure,figure*]{hypcap}
\usepackage{graphics,graphicx}
\usepackage{hyperref}
\usepackage{natbib}
\usepackage[caption=false]{subfig} % for subfloat
\usepackage{enumitem} % for specific spacing of enumerate
\usepackage{epigraph}

@@

\begin{abstract}
  The Gaia satellite is revitalizing our understanding of nearby
  open clusters and moving groups.  Here, we focus on the
  underappreciated \cn.  Based on rotation periods from TESS and
  lithium measurements from ground-based spectrographs, we find the
  age of the cluster to be $30\pm XX$\,Myr.  \sn\ is a binary system
  in the cluster, serendipitously observed by the Kepler satellite
  because the primary is nearby and Sun-like.  Kepler\,1627A was
  previously found to host a $3.56\pm 0.04\,R_\oplus$ mini-Neptune on a
  7.2\,day orbit.  We re-validate the existence of \pn, cementing it
  as the youngest planet with a well-measured age observed by the main
  Kepler mission.  Newly derived ages from Gaia offer the opportunity
  to significantly expand the census of age-dated planets -- we offer
  a literature compilation of young stars to enable this expansion.
  The properties of \pn\ are may also help clarify how the orbits and
  atmospheres of the mini-Neptune planets evolve.
\end{abstract}

@@

\section{Introduction}

At the time of the main Kepler
mission (2009--2013), only four open clusters were known in the Kepler
field: NGC\,6866, NGC\,6811, NGC\,6819,
and NGC\,6791, with ages spanning 0.7\,Gyr to 9\,Gyr
\citep{meibom_kepler_2011}.

Section~\ref{sec:first}.  Section~\ref{sec:second}.  In
Section~\ref{sec:discussion}, we discuss.
Section~\ref{sec:conclusion} gives our conclusions.

\section{first}
\label{sec:first}

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.99\textwidth]{f2.pdf}
	\end{center}
	\vspace{-0.7cm}
  \caption{
  {\bf Kepler and TESS views of the $\delta$\,Lyr cluster.} Colored
  points are kinematically selected members of the $\delta$\,Lyr
  cluster (black points in Figure~\ref{fig:XYZvtang}).  Both Kepler
  (gray panels) and TESS (colored points) observed portions of the
  cluster.  Naked-eye stars ($m_{\rm V}<6.5$) are shown in gray; four
  of them (orange points) have their names shown.  Kepler\,1627 (green
  star) was observed during the entirety of the Kepler mission, and
  has been observed  by TESS for two lunar months to date.
  \label{fig:skychart}
  }
\end{figure*}

\section{second}
\label{sec:second}

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=1\textwidth]{f4.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
    {\bf \sn\ light curve over 100 days.}  
    The full Kepler dataset spans 1{,}437 days (3.9 years), sampled at
    30 minute cadence.  The {\it top panel} shows the
    \texttt{PDCSAP} median-subtracted flux in units of
    parts-per-thousand ($\times 10^{-3}$).  The dominant signal
    is induced by starspots and plages.  The model for the
    stellar variability (orange line) is subtracted in the {\it
    bottom panel}, revealing the transits of \pn, as well as other
    deviations from the stellar variability model.  The 
    Figure Set available online shows the entire 3.9 years of
    available data.
		\label{fig:lightcurve}
	}
\end{figure*}

@@

\section{Discussion}
\label{sec:discussion}

\section{Conclusion}
\label{sec:conclusion}

@@

\section{Bonus figures that wont be included}

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=1\textwidth]{dontinclude0.png}
	\end{center}
	\vspace{-0.7cm}
	\caption{
    {\bf corner plot of gptransit fit}
	}
\end{figure*}

>>>>

\documentclass[12pt,modern,tighten]{aastex63}
%\documentclass[12pt,twocolumn,tighten,linenumbers]{aastex63}
%\documentclass[12pt,twocolumn,tighten,trackchanges]{aastex63}
\usepackage{amsmath,amstext,amssymb}
\usepackage[T1]{fontenc}
\usepackage{apjfonts}
\usepackage[figure,figure*]{hypcap}
\usepackage{graphics,graphicx}
\usepackage{hyperref}
\usepackage{natbib}
\usepackage[caption=false]{subfig} % for subfloat
\usepackage{enumitem} % for specific spacing of enumerate
\usepackage{epigraph}

@@

\begin{abstract}
  Kepler\,1627 is an active Sun-like star previously found to host a
  $3.56\pm 0.04\,R_\oplus$ mini-Neptune on a 7.2\,day orbit.  The star was
  selected for observation by the Kepler satellite because it is
  nearby and Sun-like.  Here, combining data from the Gaia satellite
  with rotation periods from TESS, we show that Kepler\,1627 is a
  member of the underappreciated \cn.  Based on isochronal and
  gyrochronal evidence, the cluster is $30\pm
  XX$\,Myr old.  Based on archival NIRC2 imaging, we also highlight the existence of 
  a previously unknown stellar companion. Our analysis advances Kepler\,1627Ab as the youngest planet with a
  precise age observed by the main Kepler mission.  Newly
  derived ages from Gaia offer the opportunity to significantly expand
  the census of age-dated planets -- we offer a literature compilation
  of young stars to support this effort.  The properties of \pn\
  may also help clarify how the orbits and atmospheres of the
  mini-Neptune planets evolve.
\end{abstract}

@@

\section{Manuscript}

While thousands of exoplanets have been discovered orbiting nearby
stars, the vast majority of them are several billion years old.  This
results in difficulty testing theories for the origins of the
different families of planets, since many expected evolutionary
processes operate on timescales of less than 100 million years.

For instance, the ``mini-Neptune'' planets, thought to be made of
rocky cores and extended atmospheric envelopes of hydrogen and helium,
are expected to shrink in size by factors of several over their first
$10^8$ years \citep{owen_constraining_2020}.  Specifically, they start
with sizes of 4--12\,$R_\oplus$ at the time of disk dispersal
($\approx$$10^7$\,years), and shrink to sizes of 2--4\,$R_\oplus$ by
10$^8$ years.  This change is driven by a combination of stellar
irradiation and internal heat that powers an outflow, which eventually
depletes or entirely strips the envelope from the rocky core
\citep{Owen_Wu_2013,gupta_sculpting_2019}.  Discovering young planets
and measuring their compositions and atmospheric outflows are key
steps toward testing this paradigm, which is often invoked to explain
the observed radius distribution of both mature exoplanets
\citep{Fulton_et_al_2017}, as well as younger planets
\citep{bouma_cluster_2020}.

% The orbits of the planets can also change early on, and can become
% significantly misaligned from the equators of their host stars.
% Measurements of the sky-projected angle between the planet orbit and
% the stellar spin (the stellar obliquity) have shown that high
% obliquities characterize hot field stars hosting both hot Jupiters and
% smaller planets \citep{winn10,louden21}.  Conversely, the obliquities
% of planets around cool stars tend to be low, but this could be due to
% tidal realignment \citep[{\it e.g.},][]{anderson21}.  Obliquity
% measurements for young planets are therefore important, because they
% provide an unobstructed view of the primordial obliquity distribution.

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.99\textwidth]{f2.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
		{\bf Kepler and TESS views of the $\delta$\,Lyr cluster.} Colored
		points are kinematically selected members of the $\delta$\,Lyr
		cluster (black points in Figure~\ref{fig:XYZvtang}).  Both Kepler
		(gray squares) and TESS (colored points) observed portions of the
		cluster.  Naked-eye stars ($m_{\rm V}<6.5$) are shown in gray; four
		of them (orange points) have their names annotated.  Kepler\,1627 (green
		star) was observed during the entirety of the Kepler mission, and
		has been observed  by TESS for two lunar months to date.
		\label{fig:skychart}
	}
\end{figure*}

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=1\textwidth]{f4.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
		{\bf 100 day window of the \sn\ light curve.}  
		The full Kepler dataset spans 1{,}437 days (3.9 years), sampled at
		30 minute cadence.  The {\it top panel} shows the
		\texttt{PDCSAP} median-subtracted flux in units of
		parts-per-thousand ($\times 10^{-3}$).  The dominant signal
		is induced by starspots and plages.  The model for the
		stellar variability (orange line) is subtracted in the {\it
			bottom panel}, revealing the transits of \pn, as well as other
		deviations from the stellar variability model.  The 
		Figure Set available online shows the entire 3.9 years of
		available data.
		\label{fig:lightcurve}
	}
\end{figure*}

@@

At the time of the main Kepler mission (2009--2013), only four open
clusters were known in the Kepler field: NGC\,6866, NGC\,6811,
NGC\,6819, and NGC\,6791, with ages spanning 0.7\,Gyr to 9\,Gyr
\citep{meibom_kepler_2011}.  Since that time, analyses of the
kinematic, photometric, and astrometric Gaia data have expanded our
knowledge of open cluster and moving group memberships \citep[{\it
e.g.},][]{cantatgaudin_gaia_2018,zari_3d_2018,kounkel_untangling_2019,Meingast2021}.
As part of our Cluster Difference Imaging Photometric Survey (CDIPS,
\citealt{bouma_cdipsI_2019}), we concatenated the available analyses
from the literature, which yielded a list of candidate young and
age-dated stars (see Section~\ref{app:targetlist}).

Comparing our young star list against the Kepler field yielded two
notable discoveries.  The first, to be discussed in an upcoming analysis
by J. Curtis et al{.}, is that Kepler observed the $\approx$350\,Myr
open cluster Theia\,520 (UBC\,1).  At least five Kepler planets are
known in the cluster, including the Kepler-52 and Kepler-968 systems.
The second discovery, and the focus of this work, is that Kepler
observed a portion of the $\delta$\,Lyr cluster, shown in
Figure~\ref{fig:skychart}.

While we have yet to perform an in-depth search of the {\bf XXX} light
curves Kepler acquired for the \cn, there is already a known planet
that has been detected in the cluster: Kepler\,1627b.  A portion of
the light curve is shown in Figure~\ref{fig:lightcurve}, and the
phase-folded transit is shown in Figure~\ref{fig:phasefold}.  The
dominant signal in the \texttt{PDCSAP} photometry is a quasi-period
starspot signal with a peak-to-peak amplitude that varies between
roughly 2\% and 8\%, depending on the degree of asymmetry in spot
coverage between the stellar hemispheres.  The starspot signal repeats
roughly every 2.606\,days, though this value varies by a few percent
depending on which portion of the light curve is analyzed {\bf (FIXME:
	do!)}.  The details of how we model this signal to derive the
parameters of the planet are discussed in Section~\ref{app:gptransit}.
The results, given in Table~\ref{tab:posterior}, are consistent with a
mini-Neptune sized planet ($3.56\pm 0.04\,R_\oplus$) on a circular
7.2\,day orbit, around a G8V star ($0.91 \pm 0.05 R_\odot$) {\bf
	fixme: age}.

Could the transit signal be produced by anything other than a planet
orbiting this near-solar analog?  \citet{morton_false_2016} considered
the various false positive scenarios, including those of a foreground
eclipsing binary, a hierarchical triple-star binary system, and a
background eclipsing binary.  \citet{morton_false_2016} reported the latter
scenario to be the only possible case, with a (model-depending)
probability of $\approx10^{-5}$.  While this statistical validation is
reassuring, ``validated" planets have been demonstrated to be false
positives \citep{shporer_three_2017}.  We are not particularly
concerned in this case however, since these cases almost always
originate when attempting to separate the smallest stars from
identically-sized gas giant planets.  With a size of $\approx
3.6\,R_\oplus$, Kepler\,1627b is well below the regime of concern.

Nonetheless, our analysis of archival Keck-NIRC2 Kp-band (2.12\,$\mu
$m) imaging did reveal the existence of a neighboring star, unresolved
in the Gaia source catalog {\bf FIXME: uncertainties}.  The NIRC2
images yield a projected separation $\rho=0\farcs17$, with $\Delta
{\rm Kp} = 2.5$.  Using the measured Gaia EDR3 parallax for the
system, this implies a projected separation of 54\,AU.  The presence
of this star is consistent with the excess noise reported to be
present in the Gaia astrometric time-series (see
Section~\ref{app:companionstar}).  Given the low chance of a star
imaged within {\bf X.X arcseconds} to be a chance unbound companion
along this line of sight, we dub the companion star Kepler\,1627B, and
proceed under the assumption that it is bound.

Unforuntately, we do not have any reliable color information about the
secondary star.  Based on the
tabulation\footnote{\url{http://www.pas.rochester.edu/~emamajek/EEM_dwarf_UBVIJHK_colors_Teff.txt},
version \texttt{2021/03/02}.} by \citet{pecaut_mamajek_2013}, the
NIR-contrast measured for a main sequence G8V star corresponds to a
spectral type for the companion of $\approx$M2V ($M_\star \approx
0.44\,M_\odot$).  However, the companion should therefore have a
significantly longer pre-main-sequence contraction phase, which would
imply that this mass is an overestimate by $\approx20\%$.

Could the companion host the planet, or be creating a false positive
signal?  The star contributes $\approx$1\% of the total flux observed
in the Kepler aperture(0.7-2.0\% V-band to G-band dmag difference from
EEM table; todo is fix using model spectra).  The observed transit has
a depth of 0.18\%.  An 18\% deep eclipse of the secondary star could
produce such a signal.  Given the shape of the observed signal
(allowed impact parameters span 0.02 to 0.73, see
Table~\ref{tab:posterior}), the body transiting the secondary would
need to be non-grazing with $R/R_2 \approx 0.42$, {\it i.e.}, with
radius $\approx 0.2R_\odot$.
The duration of the transit 
($2.823 \pm 0.057 hr$) is ultimately what makes this scenario
implausible.

% {\bf TODO:
% uhhh what about the pre-MS M dwarf star being not very dense?
% What is the density of the G dwarf expected to be?
% }
% 
% $T_{14}$ & hr & -- & 2.822 & 2.823 & 0.057 & 2.710 & 2.917 & 1086.008 & 0.002 \\
% $\rho_\star$ & g$\ $cm$^{-3}$ & -- & 2.27 & 2.312 & 0.516 & 1.408 & 3.272 & 910.746 & 0.001 \\

%\section{Discussion}
%\label{sec:discussion}

%\section{Conclusion}
%\label{sec:conclusion}

@@

\section{Transit and Stellar Variability Model}
\label{app:gptransit}

\section{Companion Star}
\label{app:companionstar}

{\bf TODO: include the RUWE vs Rp mag.}

% \section{Bonus figures that wont be included}
% 
% \begin{figure*}[t]
% 	\begin{center}
% 		\leavevmode
% 		\includegraphics[width=1\textwidth]{dontinclude0.png}
% 	\end{center}
% 	\vspace{-0.7cm}
% 	\caption{
%     {\bf corner plot of gptransit fit}
% 	}
% \end{figure*}
